---
title: "Kaggle"
author: "Shune Kawaoto"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepping the data
```{r}
# test <-  read.csv('TestSAData2NoY.csv')
# train <- read.csv('TrainSAData2.csv')
# 
# test <- test[, -1]
# train <- train[, -1]
# 
# train$Alcoholic.Status <- as.factor(train$Alcoholic.Status)
# train$sex <- as.factor(train$sex)
# train$hear_left <- as.factor(train$hear_left)
# train$hear_right <- as.factor(train$hear_right)
# train$urine_protein <- as.factor(train$urine_protein)
# train$BMI.Category <- as.factor(train$BMI.Category)
# train$AGE.Category <- as.factor(train$AGE.Category)
# train$Smoking.Status <- as.factor(train$Smoking.Status)
# 
# test$sex <- as.factor(test$sex)
# test$hear_left <- as.factor(test$hear_left)
# test$hear_right <- as.factor(test$hear_right)
# test$urine_protein <- as.factor(test$urine_protein)
# test$BMI.Category <- as.factor(test$BMI.Category)
# test$AGE.Category <- as.factor(test$AGE.Category)
# test$Smoking.Status <- as.factor(test$Smoking.Status)

completedData <- read.csv("completedData.csv")

completedData$Alcoholic.Status <- as.factor(completedData$Alcoholic.Status)
completedData$sex <- as.factor(completedData$sex)
completedData$hear_left <- as.factor(completedData$hear_left)
completedData$hear_right <- as.factor(completedData$hear_right)
# completedData$urine_protein <- as.factor(completedData$urine_protein)
completedData$BMI.Category <- as.factor(completedData$BMI.Category)
completedData$AGE.Category <- as.factor(completedData$AGE.Category)
completedData$Smoking.Status <- as.factor(completedData$Smoking.Status)

full_test <- read.csv("full_test.csv")

full_test$sex <- as.factor(full_test$sex)
full_test$hear_left <- as.factor(full_test$hear_left)
full_test$hear_right <- as.factor(full_test$hear_right)
full_test$urine_protein <- as.factor(full_test$urine_protein)
full_test$BMI.Category <- as.factor(full_test$BMI.Category)
full_test$AGE.Category <- as.factor(full_test$AGE.Category)
full_test$Smoking.Status <- as.factor(full_test$Smoking.Status)

# install.packages("Rtools")
# install.packages("pan")
# install.packages("mice")
# library(mice)

install.packages("randomForest")
library(randomForest)
# library(ggplot2)
install.packages("caret")
library(caret)
install.packages("glmnet")
library(glmnet)
install.packages("factoextra")
# library(factoextra)
library(e1071)
install.packages("xgboost")
library(xgboost)

```



## Imputation - mice (commented out)
```{r}
# imputedData <- mice(train, method='pmm', m=5, maxit=50, seed=500)
# completedData <- complete(imputedData, 1)

# write.csv(completedData, "C:\\Users\\kawao\\Documents\\UCLA\\2023 Fall\\Stats 101C\\Kaggle Project\\completedData.csv", row.names=FALSE)


# imputed_test <- mice(test, method='pmm', m=5, maxit=50, seed=500)
# full_test <- complete(imputed_test, 1)

# write.csv(full_test, "C:\\Users\\kawao\\Documents\\UCLA\\2023 Fall\\Stats 101C\\Kaggle Project\\full_test.csv", row.names=FALSE)
```



## Variable selection - LASSO
```{r}
X <- model.matrix(~ . -Alcoholic.Status -1, data = completedData)
y <- completedData$Alcoholic.Status
X <- scale(X)

set.seed(123)
cv_fit <- cv.glmnet(X, y, family = "binomial", alpha = 1)
best_lambda <- cv_fit$lambda.min
lasso_model <- glmnet(X, y, family = "binomial", alpha = 1, lambda = best_lambda)

coef(lasso_model, s = best_lambda)
```
INCLUDE: sex, age, height, weight, sight_left, SBP, DBP, BLDS, HDL_chole, LDL_chole, triglyeride, hemoglobin, urine_protein, serum_creatine, SGOT_AST, SGOT_ALT, gmma_GTP, BMI, BMI.Category, AGE, Smoking.Status

i.e. EXCLUDE: waistline, sight_right, hear_left, hear_right, and tot_chole



## Random forest - after lasso selection selection (20 predictors)
```{r}
# set.seed(123)
# rf_reduced <- randomForest(Alcoholic.Status ~ .-waistline -sight_right -hear_left -hear_right -tot_chole , data=completedData)
# output_reduced <- predict(rf_reduced, full_test)
# # kaggle score 0.72796
# 
# set.seed(123)
# rf_mtry6 <- randomForest(Alcoholic.Status ~ .-waistline -sight_right -hear_left -hear_right -tot_chole , data=completedData, mtry=6)
# output_mtry6 <- predict(rf_mtry6, full_test)
# # kaggle score 0.72646
```



## Random forest model - hanna's lasso selection (4 predictors)
```{r}
# set.seed(123)
# rf_hanna <- randomForest(Alcoholic.Status ~ sex + BMI + AGE.Category + Smoking.Status, data=completedData)
# output_hanna <- predict(rf_hanna, full_test)
# # kaggle = 0.7049
```



## Random forest model - hanna's scaled lasso (9 predictors)
```{r}
# set.seed(123)
# rf_refined <- randomForest(Alcoholic.Status~sex+age+height+DBP+HDL_chole+SGOT_AST+SGOT_ALT+gamma_GTP+Smoking.Status, data=completedData)
# output_refined <- predict(rf_refined, full_test)
# # kaggle = 0.71963
```



## Finding optimal number of predictors
```{r}
set.seed(123)
rf_full <- randomForest(Alcoholic.Status ~ ., data=completedData)
varImpPlot(rf_full)
# 
#
# set.seed(123)
# rf_reduced2 <- randomForest(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
#                            - BMI.Category - AGE.Category - weight - serum_creatinine
#                            - sight_left - sight_right - BMI, data=completedData, mtry=5)
# output_reduced2 <- predict(rf_reduced2, full_test)
# write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_reduced2), file='output_reduced2.csv', row.names = FALSE)
# # kaggle score 0.7279
# 
#
# set.seed(123)
# rf_reduced3 <- randomForest(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
#                            - BMI.Category - AGE.Category - weight - serum_creatinine
#                            - sight_left - sight_right - BMI, data=completedData)
# output_reduced3 <- predict(rf_reduced3, full_test)
# write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_reduced3), file='output_reduced3.csv', row.names = FALSE)
# # kaggle score 0.7268
# 
#
# set.seed(123)
# rf_reduced4 <- randomForest(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
#                            - BMI.Category - AGE.Category - weight - serum_creatinine
#                            - sight_left - sight_right - BMI - SGOT_AST - DBP - SBP
#                            - waistline - height - SGOT_ALT - BLDS, data=completedData, mtry=5)
# output_reduced4 <- predict(rf_reduced4, full_test)
# write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_reduced4), file='output_reduced4.csv', row.names = FALSE)
# # kaggle score 0.71296
# 
#
# set.seed(123)
# rf_reduced5 <- randomForest(Alcoholic.Status ~ gamma_GTP + hemoglobin + HDL_chole + age + sex + triglyceride, data=completedData, mtry=5)
# output_reduced5 <- predict(rf_reduced5, full_test)
# write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_reduced5), file='output_reduced5.csv', row.names = FALSE)
# kaggle score 0.6983
#
#
# set.seed(123)
# rf <- randomForest(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
#                            - BMI.Category - AGE.Category - weight - serum_creatinine
#                            - sight_left - sight_right - BMI - SGOT_AST - DBP, data=completedData, mtry=5, ntree=200)
# output_rf <- predict(rf_reduced2, full_test)
# write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_rf), file='output_rf.csv', row.names = FALSE)
# # kaggle score = 0.72763
```



## naive bayes
```{r}
# set.seed(123)
# m1 <- naiveBayes(Alcoholic.Status ~.-hear_right - hear_left - urine_protein
#                            - BMI.Category - AGE.Category - weight - serum_creatinine
#                            - sight_left - sight_right - BMI, data=completedData)
# summary(m1)
# pred_m1 <- predict(m1, full_test)
# 
# fun <- predict(rf_reduced2, completedData)
# confusionMatrix(fun, completedData$Alcoholic.Status)
# write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=pred_m1), file='naive_bayes.csv', row.names = FALSE)
# # kaggle score 0.69573
```


## Logistic
```{r}
# log_m1 <- glm(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
#                            - BMI.Category - AGE.Category - weight - serum_creatinine
#                           - sight_left - sight_right - BMI, data=completedData,
#               family=binomial())
# pred_log <- predict(log_m1, full_test)
# pred_log <- ifelse(pred_log > 0.5, "Y", "N")
# write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=pred_log), file='logistic.csv', row.names = FALSE)
# # kaggle score 0.7069
```


## optimizing mtry 3, 6, 9, 12, 15
```{r}
set.seed(123)
rf_mtry3 <- randomForest(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
                           - BMI.Category - AGE.Category - weight - serum_creatinine
                           - sight_left - sight_right - BMI, data=completedData, mtry=3)
output_mtry3 <- predict(rf_mtry3, full_test)

rf_mtry6 <- randomForest(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
                           - BMI.Category - AGE.Category - weight - serum_creatinine
                           - sight_left - sight_right - BMI, data=completedData, mtry=6)
output_mtry6 <- predict(rf_mtry6, full_test)

rf_mtry9 <- randomForest(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
                           - BMI.Category - AGE.Category - weight - serum_creatinine
                           - sight_left - sight_right - BMI, data=completedData, mtry=9)
output_mtry9 <- predict(rf_mtry9, full_test)

rf_mtry12 <- randomForest(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
                           - BMI.Category - AGE.Category - weight - serum_creatinine
                           - sight_left - sight_right - BMI, data=completedData, mtry=12)
output_mtry12 <- predict(rf_mtry12, full_test)

rf_mtry15 <- randomForest(Alcoholic.Status ~ . -hear_right - hear_left - urine_protein
                           - BMI.Category - AGE.Category - weight - serum_creatinine
                           - sight_left - sight_right - BMI, data=completedData, mtry=15)
output_mtry15 <- predict(rf_mtry15, full_test)

write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_mtry3), file='mtry3.csv', row.names = FALSE)
write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_mtry6), file='mtry6.csv', row.names = FALSE)
write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_mtry9), file='mtry9.csv', row.names = FALSE)
write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_mtry12), file='mtry12.csv', row.names = FALSE)
write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=output_mtry15), file='mtry15.csv', row.names = FALSE)

# 12 performed best --> kaggle score 0.72766
```



## try boosting next
```{r}
set.seed(123)
xgb_mice <- train(Alcoholic.Status ~ ., data=new_train, method="xgbTree",
                trControl = trainControl("cv", number=10))
pred_xgb3 <- predict(xgb_mice, full_test)
write.csv(data.frame("ID"=c(1:30000),"Alcoholic.Status"=pred_xgb3), file='pred_xgb3.csv', row.names = FALSE)
# kaggle score 0.73106
```

